\documentclass[12pt, titlepage]{article}

\usepackage{fullpage}
\usepackage{booktabs}
\usepackage{hyperref}

\usepackage[usenames, dvipsnames]{color}

\title{SFWR 3XA3: Problem Statement}
\author{Group 34 (L03) \\ Team Send Help
\\ Lau, Harrison (lauh3) \\ Mo, George (moz) \\ Truong, Vanessa (truonv1)
\\
\\ Professor: Dr. Asghar Bokhari}
\date{Due: September 22 2017}

\begin{document}

\maketitle
\newpage


\tableofcontents
\newpage




\section {Overview}
The gaming industry has rapidly grown in a matter of a few years, and continues to exponentially evolve. From the endless upgrades of video game consoles to the establishment of VR technology in modern-day gaming, the learning curve for new gamers has definitely gotten steeper. However, in the absence of all the modern, fancy embellishments, we realize that the appreciation for old-school games is still very alive. In the midst of contemporary gamers, there are still those individuals who value a simplistic and straightforward gaming experience in contrast to the very complex ones that have engulfed our gaming community today. \textbf{In response to this, we've decided to take a traditional video game, analyse its monotonous aspects, and add a modern flare to its game play.}

\section {The Problem}
The problem we seek to tackle is that which surrounds the classical game of Snake.\textbf{Though nothing is inherently wrong with the game itself, it isn't unreasonable to say that there could be more to it.} We wish to reimplement Snake in such a way that not only allows for enjoyment of the classical game in its vanilla implementation, but include an alternate version that challenges players further. By adding a little intricacy to an otherwise simple game, its entertainment value and its appeal is estimated to increase significantly.

\section {Our Goal}
Snake is no longer a game that can compete with the current mainstream games. In order to bring exposure back to a classic game, a breath of fresh air is needed to brush off the dust. \textbf{With our implementation of Snake, we hope to reignite the interest for classic games and showcase how a relatively simple game can still entertain modern gamers with slight revamping.} It is a game that can reach out to entrance-level gamers whom are transitioning into the gaming community, or even more experienced gamers who want to take a break from the complexities of advanced gaming. Overall, the game can have a flexible array of users, ranging from preschoolers and toddlers to employers on their lunch breaks.

\section {Implementation}
\textbf{Since most devices are capable of running Java or at least most people have access to a device that may run Java, our implementation will be done in Java}. This places it in an environment that is accessible to users to download and play without much effort. People affected by this would include developers who could be inspired to reinvent other classic games and the users searching to jump back into the nostalgia surrounding classic games. \textcolor{red}{The initial implementation shall be for Windows PCs.}
\end{document}
